Davos: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-09-19 paper/main.tex
minor updates to projects section

Thus, rather than constructing fully separate Python environments from scratch, 
%Davos projects serve to fill in the gap between the user's existing environment and the dependencies of their corresponding notebooks. 
Davos projects work by supplementing the user's existing environment with any additional packages (or specific package versions) needed to satisfy the dependencies of their corresponding notebooks.
In some cases, this might include every package smuggled into a notebook (e.g., if the notebook is run inside a freshly created, empty virtual environment). 
In other cases, the user's environment may already provide all required packages, and the notebook's project directory will go unused (in which case it will be deleted automatically when the notebook kernel is shut down). 
But regardless of the extent to which the existing environment is augmented, Davos's project system ensures that all smuggled packages are installed locally and loaded successfully at runtime, while the contents of the user's Python environment are never altered. 
Additionally, because \texttt{smuggle} statements in a given notebook are evaluated every time it is run, this system also ensures that the notebook's requirements will remain satisfied even if the user's Python environment changes. 
For example, suppose a user has \texttt{NumPy}~\cite{HarrEtal20} v1.24.3 installed in their current Python environment and runs a Davos-enhanced notebook that smuggles \texttt{NumPy} with ``\texttt{numpy==1.24.3}'' specified in an onion comment (see Sec.~\ref{subsec:onion}). 
Since the user's existing version of the package satisfies this requirement, Davos will happily load it into the notebook. 
But if the user later upgrades their environment's \texttt{NumPy} version to v1.25.0 (perhaps as a result of installing a different package that depends on it) and subsequently reruns this notebook, the local version will longer satisfy this requirement, so Davos will install \texttt{NumPy} v1.24.3 into the notebook's project directory and load that version instead. 
From then on, any further changes to the user's \texttt{Numpy} installation would have no effect on Davos's behavior in this particular notebook, as a satisfactory version now exists in its project directory. 
(If the version specified in the onion comment were changed, Davos would update the version installed in the project directory accordingly.) 
For efficiency, Davos projects will generally not duplicate dependencies already satisfied by the user's Python environment. 
However, if desired, adding \texttt{pip}'s \texttt{--ignore-installed} flag to an onion comment in the notebook will cause Davos to install the smuggled package into the project directory whether or not it already exists locally.

By default, each Davos-enhanced notebook will create and use its own notebook-specific project named for the absolute path to the notebook file. 
However, before smuggling its required packages, a notebook may be set to instead use an arbitrarily named, notebook-agnostic project by assigning any string (that is a valid Unix directory name) to \texttt{davos.project} (see Sec.~\ref{subsec:config}).
This provides a convenient way for multiple related notebooks that share a common set of requirements to use the same Davos project, by setting \texttt{davos.project} to the same string in each one. 
%This provides a convenient way to share a single Davos project across multiple related notebooks that share a common set of requirements, by setting \texttt{davos.project} to the same string in each one. 
It is also possible (though not recommended) to disable Davos's project system entirely and install smuggled packages directly into the user's Python environment by setting \texttt{davos.project} to \texttt{None}.

When accessed (unless its value has been set to \texttt{None}), \texttt{davos.project} will return a \texttt{Project} object that represents the project used by the current notebook (strings assigned to \texttt{davos.project} are converted to \texttt{Project}s internally). 
This object supports methods for interacting with the current project, including locating its directory on the file system, listing all installed packages' names and versions, changing the project's name, and deleting its contents altogether.
\texttt{Project} instances can also be created and managed programmatically, and Davos provides additional top-level functions for viewing and working with all existing projects (see Sec.~\ref{subsec:toplevel}).
\textcolor{red}{\textbf{TODO}: mention \texttt{AbstractProject}s? Note default project name if imported in IPython shell, or fail to get notebook name?}

>>>>

Thus, rather than constructing fully separate Python environments from scratch,
%Davos projects serve to fill in the gap between the user's existing environment and the dependencies of their corresponding notebooks.
Davos projects work by supplementing the user's existing environment with any additional packages (or specific package versions) needed to satisfy the dependencies of their corresponding notebooks.
In some cases, this might include every package smuggled into a notebook (e.g., if the notebook is run inside a freshly created, empty virtual environment).
In other cases, the user's environment may already provide all required packages, and the notebook's project directory will go unused (in which case it will be deleted automatically when the notebook kernel is shut down).
But regardless of the extent to which the existing environment is augmented, Davos's project system ensures that all smuggled packages are installed locally and loaded successfully at runtime, while the contents of the user's Python environment are never altered.

Additionally, because \texttt{smuggle} statements in a given notebook are evaluated every time it is run, this system also ensures that the notebook's requirements will remain satisfied even if the user's Python environment changes.
For example, suppose a user has \texttt{NumPy}~\cite{HarrEtal20} v1.24.3 installed in their current Python environment and runs a Davos-enhanced notebook that smuggles \texttt{NumPy} with ``\texttt{numpy==1.24.3}'' specified in an onion comment (see Sec.~\ref{subsec:onion}).
Since the user's existing version of the package satisfies this requirement, Davos will happily load it into the notebook.
But if the user later upgrades their environment's \texttt{NumPy} version to v1.25.0 (perhaps as a result of installing a different package that depends on it) and subsequently reruns this notebook, the local version will longer satisfy this requirement, so Davos will install \texttt{NumPy} v1.24.3 into the notebook's project directory and load that version instead.
From then on, any further changes to the user's \texttt{Numpy} installation would have no effect on Davos's behavior in this particular notebook, as a satisfactory version now exists in its project directory.
(If the version specified in the onion comment were changed, Davos would update the version installed in the project directory accordingly.)
For efficiency, Davos projects will generally not duplicate dependencies already satisfied by the user's Python environment.
However, if desired, adding \texttt{pip}'s \texttt{--ignore-installed} flag to an onion comment in the notebook will cause Davos to install the smuggled package into the project directory whether or not it already exists locally.

By default, each Davos-enhanced notebook will create and use its own notebook-specific project named for the absolute path to the notebook file.
However, before smuggling its required packages, a notebook may be set to instead use an arbitrarily named, notebook-agnostic project by assigning any (non-empty) string to \texttt{davos.project} (see Sec.~\ref{subsec:config}).
This provides a convenient way for multiple related notebooks that share a common set of requirements to use the same Davos project, by setting \texttt{davos.project} to the same string in each one.
%This provides a convenient way to share a single Davos project across multiple related notebooks that share a common set of requirements, by setting \texttt{davos.project} to the same string in each one.
It is also possible (though not recommended) to disable Davos's project system entirely and install smuggled packages directly into the user's Python environment by setting \texttt{davos.project} to \texttt{None}.

When accessed (unless its value has been set to \texttt{None}), \texttt{davos.project} will return a \texttt{Project} object that represents the project used by the current notebook (strings assigned to \texttt{davos.project} are converted to \texttt{Project}s internally). This object supports methods for interacting with the current project, including locating its directory on the file system, listing all installed packages' names and versions, changing the project's name, and deleting its contents altogether.
\texttt{Project} instances can also be created and managed programmatically, and Davos provides additional utilities for viewing and working with all existing projects (see Secs.~\ref{subsec:config} and \ref{subsec:toplevel}).
\textcolor{red}{\textbf{TODO}: mention \texttt{AbstractProject}s? Note default project name if imported in IPython shell, or fail to get notebook name?}
